Ethereum: A Secure Decentralised Generalised Transaction Ledger

### 2014-04-07 Paper.tex
Uncles and what not.

\section{Proposal} \label{ch:proposal}

\subsection{The Blockchain Paradigm} \label{ch:overview}

@@

\begin{eqnarray}
\mathbb{S}_{t+1} & \equiv & \Pi(\mathbb{S}_t, \mathbf{T}) \\
\mathbf{T} & \equiv & [ T_0, T_1, ... ] \\
\Pi(\mathbb{S}, \mathbf{T}) & \equiv & \Omega(\Upsilon(\Upsilon(\mathbb{S}, T_0), T_1) ...)
\end{eqnarray}

Where $\Omega$ is the block-finalisation state transition function (a function that rewards a nominated party); $\mathbf{T}$ is the series of transactions included in this block; and $\Pi$ is the block-level state-transition function.

@@

\subsubsection{Value}

@@

\subsubsection{Which History?}

@@

\subsection{Blocks, State and Transactions} \label{ch:bst}

@@

\subsubsection{World State} \label{ch:state}

@@

\subsubsection{Transaction} \label{ch:transaction}

@@

\subsubsection{Block} \label{ch:block}

The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with a set of transactions and a set of other block headers that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{uncles}). The block header contains several pieces of information:

\begin{description}
\item[parentHash] The SHA3 256-bit hash of the parent block, in its entirety; formally $H_p$.
\item[unclesHash] The SHA3 256-bit hash of the uncles list portion of this block; formally $H_u$.
\item[coinbase] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_b$.
\item[stateRoot] The SHA3 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
\item[transactionsHash] The SHA3 256-bit hash of the transactions list portion of this block; formally $H_t$.
\item[difficulty] A scalar value corresonding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
\item[extraData] An arbitrary byte array containing data relavant to this block. Generally 256-bit or less, but could be much larger in the genesis block; formally $H_x$.
\item[nonce] A 256-bit hash which proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
\end{description}

The difficulty of a block $H$, is defined as $H_d$:

\begin{equation}
H_d \equiv  \begin{cases}
2^{22} & \quad \text{if genesis block}\\
H'_d + \frac{H'_d}{1024} & \quad \text{if $H_t < H'_t + 42$}\\
H'_d - \frac{H'_d}{1024} & \quad \text{otherwise}\\
\end{cases}
\end{equation}

Where $H'$ is the previous block, $H_t$ is the timestamp of block $H$.

@@

The nonce, $H_n$, must satisfy the relation:

\begin{equation}
\mathcal{P}(H, H_n) \leqslant \frac{2^{256}}{H_d}
\end{equation}

@@

This is the foundation of the security of the blockchain and is the fundamental reason why a malicious node cannot propogate newly created blocks that would otherwise overwrite (``rewrite'') history. Because the nonce must satisfy this requirement, and because satisfaction of this requirement both depends on the contents of the block and

The other two components in the block are simply a list of the transactions (as defined previously) and a list of uncle block headers (of the same format as above).

\subsection{Gas and Payment} \label{ch:payment}

@@

\subsection{Transaction Execution} \label{ch:transactions}

@@

\begin{enumerate}
\item The transaction signature is valid;
\item the transaction nonce is valid (equivalent to the sender account's current nonce);
\item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction;
\item the sender account balance contains at least the cost, $v_0$, (in wei) required in up-front payment;
\end{enumerate}

@@

@@

After the message call or contract creation is processed, the state is finalised by refunding $g'$, the remaining gas, to the sender at the original rate. So we define the final state $\mathbb{S}'$ in terms of the provisional state $\mathbb{S}_P$.

\begin{eqnarray}
\mathbb{S}' & \equiv & \mathbb{S}_o \quad \text{except} \\
\mathbb{S}'[s]_{balance} & \equiv & \mathbb{S}_o[s]_{balance} + g' T_p
\end{eqnarray}

\subsection{Contract Creation} \label{ch:create}

@@

\begin{eqnarray}
A(s, o) & \equiv & a \quad \text{where:} \\
a' & \equiv & \mathcal{B}_{96..255}(\mathcal{S}(\mathcal{R}([s, o]))) \\
a & \equiv & \arg \min_x : x \geqslant a' \wedge \mathbb{S}_0[x] = \varnothing
\end{eqnarray}

@@

\subsection{Message Call} \label{ch:call}

@@

\subsection{Execution Model} \label{ch:model}

@@

\subsubsection{Basics}

@@

\subsubsection{Fees Overview}

@@

\subsubsection{Execution Environment}

@@

\subsubsection{Execution Overview}

@@

\subsubsection{The Execution Cycle}

@@

The gas is reduced by the instruction's gas cost and for most instructions, the program counter increments on each cycle, for the three exceptions, we assume a function $J$ which evaluates to the according value:
\begin{eqnarray}
\mathbb{M}'_{g} & \equiv & \mathbb{M}_{g} - C(\mathbb{S}, \mathbb{M}, I) \\
\mathbb{M}'_{pc} & \equiv & \begin{cases}
J_o(\mathbb{M}) & \text{if} \quad w \in \{ \text{\small JUMP, JUMPI} \} \\
\mathbb{M}_{pc} + p & \text{if} \quad w \in [\text{\small PUSH1}, \text{\small PUSH32}] \\
\mathbb{M}_{pc} + 1 & \text{otherwise}
\end{cases}
\end{eqnarray}

@@

\subsection{Blocktree to Blockchain} \label{ch:ghost}

@@

\begin{eqnarray}
B_{td} & \equiv & P_{td} + B_d + \sum\limits_{U \in B_\mathbb{U}} U_d \\
P & \equiv & B_P
\end{eqnarray}

As such given a block $B$, $B_{td}$ is its total difficulty, $B_P$ is its parent block, $B_d$ is its difficulty and $B_\mathbb{U}$ is its set of uncle blocks.

\subsection{Block Finalisation} \label{ch:finalisation}

Validate uncles.

Reward everything.

Compute proof-of-work.

@@

\section{User Interface}

Browser-style.

Displays ETH balance (i.e. functions as a wallet).

Tabs or perhaps MacOS style app icons.

Display taken up with a single app.

Extensible \eg should be able to display the balances of other currencies hosted on Etheruem alongside ETH.

Should be able to trust certain apps with creation of transactions (and thus sending of ETH \& payment of gas).

Preconfigured with home app that is the Eth app store.

Eth app store comes preloaded with high trust for ethereum.org; apps have identity and WoT reputation.

@@

@@

TODO: Specify minimum memory limits for a general cost definition.

@@

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{30s: Environmental Information}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x30 & {\small ADDRESS} & 0 & 1 & Get address of currently executing account. \\
&&&& $\mathbb{M}'_T[0] \equiv I_a$ \\
\midrule
0x31 & {\small BALANCE} & 0 & 1 & Get balance of currently executing account. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{S}[I_a]_{balance}$ \\
\midrule
0x32 & {\small ORIGIN} & 0 & 1 & Get execution origination address. \\
&&&& $\mathbb{M}'_T[0] \equiv I_o$ \\
&&&& This is the sender of original transaction; it's never a contract. \\
\midrule
0x33 & {\small CALLER} & 0 & 1 & Get caller address. \\
&&&& $\mathbb{M}'_T[0] \equiv I_c$ \\
&&&& This is the address of the account that is directly responsible for this execution. \\
\midrule
0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for this execution. \\
&&&& $\mathbb{M}'_T[0] \equiv I_v$ \\
\midrule
0x35 & {\small CALLDATALOAD} & 1 & 1 & Get input data of current environment. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathcal{I}_\mathbf{d}[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 31) ]$ \\
&&&& This pertains to the input data passed with the message call instruction or transaction. \\
\midrule
0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current environment. \\
&&&& $\mathbb{M}'_T[0] \equiv \lVert \mathcal{I}_\mathbf{d} \rVert$ \\
&&&& This pertains to the input data passed with the message call instruction or transaction. \\
\midrule
0x37 & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathcal{I}_p$ \\
&&&& This is gas price specified by the originating transaction.\\
\bottomrule
\end{tabular*}

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{40s: Block Information}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x40 & {\small PREVHASH} & 0 & 1 & Get hash of most recent complete block. \\
&&&& $\mathbb{M}'_T[0] \equiv I_{H_{prevHash}}$ \\
&&&& $I_{H_{prevHash}}$ is the previous block's hash. \\
\midrule
0x41 & {\small COINBASE} & 0 & 1 & Get the block's coinbase address. \\
&&&& $\mathbb{M}'_T[0] \equiv I_{H_{coinbase}}$ \\
\midrule
0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
&&&& $\mathbb{M}'_T[0] \equiv I_{H_{timestamp}}$ \\
\midrule
0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
&&&& $\mathbb{M}'_T[0] \equiv I_{H_{number}}$ \\
&&&& $I_{H_{number}}$ is the total number of ancestor blocks of the current block. \\
\midrule
0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
&&&& $\mathbb{M}'_T[0] \equiv I_{H_{difficulty}}$ \\
\midrule
0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
&&&& $\mathbb{M}'_T[0] \equiv 10^6$ \\
&&&& Presently, the gas limit is hard-coded in the protocol. \\
\bottomrule
\end{tabular*}

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{50s: Stack, Memory, Storage and Flow Operations}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x50 & {\small POP} & 1 & 0 & Remove item from stack. \\
\midrule
0x51 & {\small DUP} & 1 & 2 & Duplicate stack item. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[0]$ \\
\midrule
0x52 & {\small SWAP} & 2 & 2 & Exchange stack items. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[1]$ \\
&&&& $\mathbb{M}'_T[1] \equiv \mathbb{M}_T[0]$ \\
\midrule
0x53 & {\small MLOAD} & 1 & 1 & Load word from memory. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_M[\mathbb{M}_T[0] \dots (\mathbb{M}_T[0] + 31) ]$ \\
\midrule
0x54 & {\small MSTORE} & 2 & 0 & Save word to memory. \\
&&&& $\mathbb{M}'_M[ \mathbb{M}_T[0] \dots (\mathbb{M}_T[0] + 31) ] \equiv \mathbb{M}_T[1]$ \\
\midrule
0x55 & {\small MSTORE8} & 2 & 0 & Save byte to memory. \\
&&&& $\mathbb{M}'_M[ \mathbb{M}_T[0] ] \equiv (\mathbb{M}_T[1] \mod 256) $ \\
\midrule
0x56 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_S[\mathbb{M}_T[0]]$ \\
\midrule
0x57 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
&&&& $\mathbb{M}'_S[ \mathbb{M}_T[0] ] \equiv \mathbb{M}_T[1] $ \\
\midrule
0x58 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
&&&& $J_{\text{\tiny JUMP}}(\mathbb{M}) \equiv \mathbb{M}_T[0] $ \\
&&&& This has the effect of writing said value to $\mathbb{M}_{pc}$. See section \ref{ch:model}. \\
\midrule
0x59 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
&&&& $J_{\text{\tiny JUMPI}}(\mathbb{M}) \equiv \begin{cases} \mathbb{M}_T[0] & \text{if} \quad \mathbb{M}_T[0] = 0 \\ \mathbb{M}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
&&&& This has the effect of writing said value to $\mathbb{M}_{pc}$. See section \ref{ch:model}. \\
\midrule
0x5a & {\small PC} & 0 & 1 & Get the program counter. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_{pc}$ \\
\midrule
0x5b & {\small MSIZE} & 0 & 1 & Get the size of active memory. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_{i}$ \\
\midrule
0x5c & {\small GAS} & 0 & 1 & Get the amount of available gas. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_{g}$ \\
\bottomrule
\end{tabular*}

>>>>

\section{The Blockchain Paradigm} \label{ch:overview}

@@

\begin{eqnarray}
\mathbb{S}_{t+1} & \equiv & \Pi(\mathbb{S}_t, \mathbf{T}) \\
B & \equiv & (..., [ T_0, T_1, ... ]) \\
\Pi(\mathbb{S}, B) & \equiv & \Omega(B, \Upsilon(\Upsilon(\mathbb{S}, T_0), T_1) ...)
\end{eqnarray}

Where $\Omega$ is the block-finalisation state transition function (a function that rewards a nominated party); $B$ is this block, which includes a series of transactions amongst some other components; and $\Pi$ is the block-level state-transition function.

@@

\subsection{Value}

@@

\subsection{Which History?}

@@

\section{Blocks, State and Transactions} \label{ch:bst}

@@

\subsection{World State} \label{ch:state}

@@

\subsection{The Transaction} \label{ch:transaction}

@@

\subsection{The Block} \label{ch:block}

The block in Ethereum is the collection of relevant pieces of information (known as the block \textit{header}), $H$, together with a set of transactions, $\mathbf{T}$, and a set of other block headers $\mathbf{U}$ that are known to have a parent equal to the present block's parent's parent (such blocks are known as \textit{uncles}). The block header contains several pieces of information:

\begin{description}
\item[parentHash] The SHA3 256-bit hash of the parent block, in its entirety; formally $H_p$.
\item[unclesHash] The SHA3 256-bit hash of the uncles list portion of this block; formally $H_u$.
\item[coinbase] The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_b$.
\item[stateRoot] The SHA3 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_r$.
\item[transactionsHash] The SHA3 256-bit hash of the transactions list portion of this block; formally $H_t$.
\item[difficulty] A scalar value corresonding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_d$.
\item[timestamp] A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally $H_s$.
\item[extraData] An arbitrary byte array containing data relavant to this block. Generally 256-bit or less, but could be much larger in the genesis block. Not formally used.
\item[nonce] A 256-bit hash which proves that a sufficient amount of computation has been carried out on this block; formally $H_n$.
\end{description}

The other two components in the block are simply a list of the transactions (as defined previously) and a list of uncle block headers (of the same format as above). Formally, we can refer to a block $B$:
\begin{equation}
B \equiv (H, \mathbf{T}, \mathbf{U})
\end{equation}

We can assert its validity if and only if it satisfies several conditions: It must be internally consistent with the uncle and transaction block hashes; the given transactions $\mathbf{T}$, when executed in order on the base state $\mathbb{S}$ result in a new state of the identity $H_r$:
\begin{eqnarray}
& & \mathcal{S}(\mathcal{R}(\mathbf{U})) \equiv H_u \\
& \wedge & \mathcal{S}(\mathcal{R}(\mathbf{T})) \equiv H_t \\
& \wedge & r(\Pi(\mathbb{S}, B)) \equiv H_r
\end{eqnarray}

where $r(\mathbb{S})$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of $\mathbb{S}$ with values encoded using RLP.

\subsubsection{Block Header Validity}

We define $P(B_H)$ to be the parent block of $B$, formally:
\begin{equation}
P(H) \equiv B': S(R(B')) = H_p
\end{equation}

The canonical difficulty of a block of header $H$, is defined as $D(H)$:
\begin{equation}
D(H) \equiv \begin{cases}
2^{22} & \quad \text{if genesis block}\\
{P(H)_H}_d + \frac{{P(H)_H}_d}{1024} & \quad \text{if $H_t < {P(H)_H}_t + 42$}\\
{P(H)_H}_d - \frac{{P(H)_H}_d}{1024} & \quad \text{otherwise}\\
\end{cases}
\end{equation}

$H_t$ is the timestamp of block $H$, must fulfill the relation:
\begin{equation}
H_t > {P(H)_H}_t
\end{equation}

@@

The nonce, $H_n$, must satisfy the relation:
\begin{equation}
\mathcal{P}(H, H_n) \leqslant \frac{2^{256}}{H_d}
\end{equation}

@@

This is the foundation of the security of the blockchain and is the fundamental reason why a malicious node cannot propogate newly created blocks that would otherwise overwrite (``rewrite'') history. Because the nonce must satisfy this requirement, and because satisfaction of this requirement depends on the contents of the block and in turn its composed transactions, creating new, valid, blocks is difficult, and, over time, requires approximately the total compute power of the trustworthy portion of the mining peers.

Thus we are able to define the block header validity function $V(H)$:
\begin{eqnarray}
V(H) & \equiv & \mathcal{P}(H, H_n) \leqslant \frac{2^{256}}{H_d} \\
& \wedge & H_d = D(H) \\
& \wedge & H_t > {P(H)_H}_t
\end{eqnarray}

\section{Gas and Payment} \label{ch:payment}

@@

\section{Transaction Execution} \label{ch:transactions}

@@

\begin{enumerate}
\item The transaction signature is valid;
\item the transaction nonce is valid (equivalent to the sender account's current nonce);
\item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction;
\item the sender account balance contains at least the cost, $v_0$, required in up-front payment;
\end{enumerate}

@@

The transaction must not require more gas to be utilised in this block than the global block limit of gas, specified at $10^6$.

@@

After the message call or contract creation is processed, the state is finalised by refunding $g'$, the remaining gas, to the sender at the original rate and the Ether for the gas actually used is given to the miner, whose address is specified as the coinbase of the present block $B$. So we define the final state $\mathbb{S}'$ in terms of the provisional state $\mathbb{S}_P$:

\begin{eqnarray}
\mathbb{S}' & \equiv & \mathbb{S}_o \quad \text{except} \\
\mathbb{S}'[s]_{balance} & \equiv & \mathbb{S}_o[s]_{balance} + g' T_p \\
\mathbb{S}'[m]_{balance} & \equiv & \mathbb{S}_o[m]_{balance} + (g - g') T_p \\
m & \equiv & {B_H}_b
\end{eqnarray}

\section{Contract Creation} \label{ch:create}

@@

\begin{eqnarray}
A(s, o) & \equiv & a \quad \text{where:} \\
a & = & \arg \min_x : x \geqslant a' \wedge \mathbb{S}_0[x] = \varnothing \\
a' & = & \mathcal{B}_{96..255}(\mathcal{S}(\mathcal{R}([s, o])))
\end{eqnarray}

@@

\section{Message Call} \label{ch:call}

@@

\section{Execution Model} \label{ch:model}

@@

\subsection{Basics}

@@

\subsection{Fees Overview}

@@

\subsection{Execution Environment}

@@

\subsection{Execution Overview}

@@

\subsection{The Execution Cycle}

@@

The gas is reduced by the instruction's gas cost and for most instructions, the program counter increments on each cycle, for the three exceptions, we assume a function $J$, subscripted by one of two instructions, which evaluates to the according value:
\begin{eqnarray}
\mathbb{M}'_{g} & \equiv & \mathbb{M}_{g} - C(\mathbb{S}, \mathbb{M}, I) \\
\mathbb{M}'_{pc} & \equiv & \begin{cases}
J_o(\mathbb{M}) & \text{if} \quad w \in \{ \text{\small JUMP, JUMPI} \} \\
\mathbb{M}_{pc} + p & \text{if} \quad w \in [\text{\small PUSH1}, \text{\small PUSH32}] \\
\mathbb{M}_{pc} + 1 & \text{otherwise}
\end{cases}
\end{eqnarray}

@@

\section{Blocktree to Blockchain} \label{ch:ghost}

@@

\begin{eqnarray}
B_t & \equiv & B'_t + B_d + \sum\limits_{U \in B_\mathbf{U}} U_d \\
B' & \equiv & P(H) \\
H & \equiv & B_H \\
\end{eqnarray}

As such given a block $B$, $B_t$ is its total difficulty, $B'$ is its parent block, $B_d$ is its difficulty and $B_\mathbf{U}$ is its set of uncle blocks.

\section{Block Finalisation} \label{ch:finalisation}

The process of finalising a block involves three stages:

\begin{enumerate}
\item Validate (or, if mining, determine) uncles.
\item Apply rewards.
\item Verify (or, if mining, compute a valid) nonce.
\end{enumerate}

\subsubsection{Uncle Validation}

The validation of uncle headers means nothing more than verifying that each uncle header is both a valid header and satisfies the relation of uncle to the present block. Formally:

\begin{equation}
\bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; P(U) = P(P(B_H)) \; \wedge \; P(B_H) \neq B
\end{equation}

\subsubsection{Reward Application}

The application of rewards to a block involves little more

@@

%\section{User Interface}
%
%Browser-style.
%Displays ETH balance (i.e. functions as a wallet).
%Tabs or perhaps MacOS style app icons.
%Display taken up with a single app.
%Extensible \eg should be able to display the balances of other currencies hosted on Etheruem alongside ETH.
%Should be able to trust certain apps with creation of transactions (and thus sending of ETH \& payment of gas).
%Preconfigured with home app that is the Eth app store.
%Eth app store comes preloaded with high trust for ethereum.org; apps have identity and WoT reputation.

@@

\section{Future Directions} \label{ch:future}

Rether than being hard-specified in the protocol, the block limit will become based upon some fixed ratio about a long-term moving average of actual gas usage in blocks, allowing it to slowly increase according to the network's ability to manage the transaction volume.

@@

The general gas cost function is defined as:

\begin{equation}
C(\mathbb{S}, \mathbb{M}) \equiv C_M(\mathbb{S}, \mathbb{M}) + \begin{cases}
c_\text{\tiny SSTORE} & \text{if} \quad o = \text{\small SSTORE} \\
G_{call} & \text{if} \quad o = \text{\small CALL}\\
G_{create} & \text{if} \quad o = \text{\small CREATE}\\
G_{sha3} & \text{if} \quad o = \text{\small SHA3}\\
G_{sload} & \text{if} \quad o = \text{\small SLOAD}\\
G_{balance} & \text{if} \quad o = \text{\small BALANCE}\\
G_{stop} & \text{if} \quad o = \text{\small STOP}\\
G_{step} & \text{otherwise}
\end{cases}
\end{equation}

where $c_o$ is specified in the appropriate section below and $C_M$ is the memory cost function, given as the product of $G_{memory}$ and the maximum of 0 and the ceiling of the number of words in size that the memory must be over the current number of words, $\mathbb{M}_i$ in order that all accesses reference valid memory whether for read or write; $\mathbb{M}'_i$ is defined as this new maximum number of words of active memory.

@@

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{30s: Environmental Information}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x30 & {\small ADDRESS} & 0 & 1 & Get address of currently executing account. \\
&&&& $\mathbb{M}'_T[0] \equiv I_a$ \\
\midrule
0x31 & {\small BALANCE} & 0 & 1 & Get balance of currently executing account. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{S}[I_a]_{balance}$ \\
\midrule
0x32 & {\small ORIGIN} & 0 & 1 & Get execution origination address. \\
&&&& $\mathbb{M}'_T[0] \equiv I_o$ \\
&&&& This is the sender of original transaction; it's never a contract. \\
\midrule
0x33 & {\small CALLER} & 0 & 1 & Get caller address. \\
&&&& $\mathbb{M}'_T[0] \equiv I_c$ \\
&&&& This is the address of the account that is directly responsible for this execution. \\
\midrule
0x34 & {\small CALLVALUE} & 0 & 1 & Get deposited value by the instruction/transaction responsible for this execution. \\
&&&& $\mathbb{M}'_T[0] \equiv I_v$ \\
\midrule
0x35 & {\small CALLDATALOAD} & 1 & 1 & Get input data of current environment. \\
&&&& $\mathbb{M}'_T[0] \equiv I_\mathbf{d}[ \mathbb{M}_S[0] \dots (\mathbb{M}_S[0] + 31) ]$ \\
&&&& This pertains to the input data passed with the message call instruction or transaction. \\
\midrule
0x36 & {\small CALLDATASIZE} & 0 & 1 & Get size of input data in current environment. \\
&&&& $\mathbb{M}'_T[0] \equiv \lVert I_\mathbf{d} \rVert$ \\
&&&& This pertains to the input data passed with the message call instruction or transaction. \\
\midrule
0x37 & {\small GASPRICE} & 0 & 1 & Get price of gas in current environment. \\
&&&& $\mathbb{M}'_T[0] \equiv I_p$ \\
&&&& This is gas price specified by the originating transaction.\\
\bottomrule
\end{tabular*}

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{40s: Block Information}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x40 & {\small PREVHASH} & 0 & 1 & Get hash of most recent complete block. \\
&&&& $\mathbb{M}'_T[0] \equiv {I_H}_p$ \\
&&&& $I_{H_p}$ is the previous block's hash. \\
\midrule
0x41 & {\small COINBASE} & 0 & 1 & Get the block's coinbase address. \\
&&&& $\mathbb{M}'_T[0] \equiv {I_H}_b$ \\
\midrule
0x42 & {\small TIMESTAMP} & 0 & 1 & Get the block's timestamp. \\
&&&& $\mathbb{M}'_T[0] \equiv {I_H}_t$ \\
\midrule
0x43 & {\small NUMBER} & 0 & 1 & Get the block's number. \\
&&&& $\mathbb{M}'_T[0] \equiv n({I_H})$ \\
&&&& $n(I_H)$ is the total number of ancestor blocks of the current block. \\
\midrule
0x44 & {\small DIFFICULTY} & 0 & 1 & Get the block's difficulty. \\
&&&& $\mathbb{M}'_T[0] \equiv {I_H}_d$ \\
\midrule
0x45 & {\small GASLIMIT} & 0 & 1 & Get the block's gas limit. \\
&&&& $\mathbb{M}'_T[0] \equiv 10^6$ \\
&&&& Presently, the gas limit is hard-coded in the protocol. \\
\bottomrule
\end{tabular*}

\begin{tabular*}{\columnwidth}[h]{rlrrl}
\toprule
\multicolumn{5}{c}{\textbf{50s: Stack, Memory, Storage and Flow Operations}} \vspace{5pt} \\
\textbf{Value} & \textbf{Mnemonic} & $\delta$ & $\alpha$ & \textbf{Description} \vspace{5pt} \\
0x50 & {\small POP} & 1 & 0 & Remove item from stack. \\
\midrule
0x51 & {\small DUP} & 1 & 2 & Duplicate stack item. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[0]$ \\
\midrule
0x52 & {\small SWAP} & 2 & 2 & Exchange stack items. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_T[1]$ \\
&&&& $\mathbb{M}'_T[1] \equiv \mathbb{M}_T[0]$ \\
\midrule
0x53 & {\small MLOAD} & 1 & 1 & Load word from memory. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_M[\mathbb{M}_T[0] \dots (\mathbb{M}_T[0] + 31) ]$ \\
\midrule
0x54 & {\small MSTORE} & 2 & 0 & Save word to memory. \\
&&&& $\mathbb{M}'_M[ \mathbb{M}_T[0] \dots (\mathbb{M}_T[0] + 31) ] \equiv \mathbb{M}_T[1]$ \\
\midrule
0x55 & {\small MSTORE8} & 2 & 0 & Save byte to memory. \\
&&&& $\mathbb{M}'_M[ \mathbb{M}_T[0] ] \equiv (\mathbb{M}_T[1] \mod 256) $ \\
\midrule
0x56 & {\small SLOAD} & 1 & 1 & Load word from storage. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{S}'[I_a]_{storage}[\mathbb{M}_T[0]]$ \\
\midrule
0x57 & {\small SSTORE} & 2 & 0 & Save word to storage. \\
&&&& $\mathbb{S}'[I_a]_{storage}[ \mathbb{M}_T[0] ] \equiv \mathbb{M}_T[1] $ \\
&&&& $c_{\text{\tiny SSTORE}} \equiv \begin{cases}
2G_{sstore} & \text{if} \quad \mathbb{M}_T[1] \neq 0 \; \wedge \; \mathbb{S}[I_a]_{storage}[\mathbb{M}_T[0]] = 0 \\
0 & \text{if} \quad \mathbb{M}_T[1] = 0 \; \wedge \; \mathbb{S}[I_a]_{storage}[\mathbb{M}_T[0]] \neq 0 \\
G_{sstore} & \text{otherwise}
\end{cases}$ \\
\midrule
0x58 & {\small JUMP} & 1 & 0 & Alter the program counter. \\
&&&& $J_{\text{\tiny JUMP}}(\mathbb{M}) \equiv \mathbb{M}_T[0] $ \\
&&&& This has the effect of writing said value to $\mathbb{M}_{pc}$. See section \ref{ch:model}. \\
\midrule
0x59 & {\small JUMPI} & 2 & 0 & Conditionally alter the program counter. \\
&&&& $J_{\text{\tiny JUMPI}}(\mathbb{M}) \equiv \begin{cases} \mathbb{M}_T[0] & \text{if} \quad \mathbb{M}_T[0] = 0 \\ \mathbb{M}_{pc} + 1 & \text{otherwise} \end{cases} $ \\
&&&& This has the effect of writing said value to $\mathbb{M}_{pc}$. See section \ref{ch:model}. \\
\midrule
0x5a & {\small PC} & 0 & 1 & Get the program counter. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_{pc}$ \\
\midrule
0x5b & {\small MSIZE} & 0 & 1 & Get the size of active memory. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_{i}$ \\
\midrule
0x5c & {\small GAS} & 0 & 1 & Get the amount of available gas. \\
&&&& $\mathbb{M}'_T[0] \equiv \mathbb{M}_{g}$ \\
\bottomrule
\end{tabular*}
